---
title: "edivalo-seedlings-trait-demo"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse) # data structuring
library(ggplot2) # plotting
library(ggthemes) # plotting
library(GGally) # plotting

source('edivalo-seedlings-traits.R', local=TRUE) # script with the trait data
theme_cw <- function () { 
  theme_bw(base_size=12) %+replace% 
    theme(
      panel.background = element_blank(), 
      plot.background = element_blank(), 
      axis.ticks = element_line(colour = "grey70", size = rel(0.5)),
      panel.grid.minor = element_blank(), 
      panel.grid.major.x = element_blank(),
      legend.background = element_blank(), 
      legend.key = element_blank(),
      strip.background = element_blank(), 
      strip.text=element_text(size=12),
      axis.text=element_text(size=12),
      complete = TRUE
    )
}

```

## Correlations between trait variables
Simple correlation plot: 1. root bm, shoot bm, and total bm are strongly correlated, will probably use the total bm 2. sla (first 3 leaves) and sla.2 (first 2 leaves only) are strongly correlated, that's good, we can just use the first 2 leaves. 
```{r corr-plot1, echo=FALSE, warning=FALSE, message=FALSE}
ggcorr(species.size.6)
```

More detailed correlation plot with our smaller set of variables
```{r corr-plot2, echo=FALSE, warning=FALSE, message=FALSE}
ggpairs(species.size.7, columns = c('bm.tot', 'len.sh', 'rt.sh.bm', 'sla.2','c.n.ratio'))

```

## Trait distributions
Seeing what the distribution of trait relationships are, and where our toothpick-marked seedlings fall in this space (these species are labled in red)

Total biomass and root-shoot biomass ratio
```{r corr-plot3, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(species.size.7, aes(x=bm.tot, y = rt.sh.bm)) + 
  geom_text(aes(label=species, color=toothpicks),hjust=0, vjust=0) +
  scale_color_manual(values=c('black','red')) +
  geom_smooth(method = 'lm', se = FALSE, color = 'grey50') + 
  theme_cw() +
  theme(legend.position = 'none')
```


SLA and C:N
```{r corr-plot4, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(species.size.7, aes(x=sla.2, y = c.n.ratio)) + 
  geom_text(aes(label=species, color=toothpicks),hjust=0, vjust=0) +
  scale_color_manual(values=c('black','red')) +
  geom_smooth(method = 'lm', se = FALSE, color = 'grey50') + 
  theme_cw() +
  theme(legend.position = 'none')
```

Total biomass and shoot length
```{r corr-plot5, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(species.size.7, aes(x=bm.tot, y = len.sh)) + 
  geom_text(aes(label=species, color=toothpicks),hjust=0, vjust=0) +
  scale_color_manual(values=c('black','red')) +
  geom_smooth(method = 'lm', se = FALSE, color = 'grey50') + 
  theme_cw() +
  theme(legend.position = 'none')
```