---
title: "eDiValo Seedlings Natural Regeneration"
output:
  word_document: default
  html_document: default
---

## Setup and data structuring
Loads libraries, sets up custom theme for ggplot, and reads in data (code not printed)
```{r setup2, include=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding

library(plyr)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(lme4)
library(lmerTest)
library(gtsummary)
source('edivalo-seedlings-traits.R', local=TRUE) # script with the trait data

# theme for plotting
theme_cw <- function () { 
  theme_bw(base_size=12) %+replace% 
    theme(
      panel.background = element_blank(), 
      plot.background = element_blank(), 
      axis.ticks = element_line(colour = "grey70", size = rel(0.5)),
      panel.grid.minor = element_blank(), 
      panel.grid.major.x = element_blank(),
      legend.background = element_blank(), 
      legend.key = element_blank(),
      strip.background = element_blank(), 
      strip.text=element_text(size=12),
      axis.text=element_text(size=12),
      complete = TRUE
    )
}

# Chhaya's path
nat.seedlings.10 <- read.csv("~/Dropbox/eDiValo-seedlings/Data Entry/natural_regen_2019_10.csv", stringsAsFactors = FALSE)
nat.seedlings.03 <- read.csv("~/Dropbox/eDiValo-seedlings/Data Entry/natural_regen_2020_03.csv", stringsAsFactors = FALSE)
nat.seedlings.05 <- read.csv("~/Dropbox/eDiValo-seedlings/Data Entry/natural_regen_2020_05.csv", stringsAsFactors = FALSE)
nat.seedlings.06 <- read.csv("~/Dropbox/eDiValo-seedlings/Data Entry/natural_regen_2020_06.csv", stringsAsFactors = FALSE)

# Want to combine these together
# remove . and spaces
names(nat.seedlings.10) <- gsub("[^a-z]","", names(nat.seedlings.10))
names(nat.seedlings.03) <- gsub("[^a-z]","", names(nat.seedlings.03)) 
names(nat.seedlings.10) <- gsub("[^a-z]","", names(nat.seedlings.10))
names(nat.seedlings.03) <- gsub("[^a-z]","", names(nat.seedlings.03)) 
names(nat.seedlings.05) <- gsub("[^a-z]","", names(nat.seedlings.05))
names(nat.seedlings.06) <- gsub("[^a-z]","", names(nat.seedlings.06))

# small cleaning issues
nat.seedlings.10$subplot <- ifelse(nat.seedlings.10$subplot==1, "O", "I")
nat.seedlings.03$falvul <- as.integer(substr(nat.seedlings.03$falvul, 1, 1))
nat.seedlings.05 <- nat.seedlings.05[,1:43] # removing blank columsn at the end

# add in year_month time point
nat.seedlings.10$month <- '2019_10'
nat.seedlings.03$month <- '2020_03'
nat.seedlings.05$month <- '2020_05'
nat.seedlings.06$month <- '2020_06'

nat.seedlings.a <- full_join(nat.seedlings.10, nat.seedlings.03)
nat.seedlings.b <- full_join(nat.seedlings.05, nat.seedlings.06)
nat.seedlings <- full_join(nat.seedlings.a, nat.seedlings.b)
nat.seedlings <- nat.seedlings %>% select(month, everything()) # move month to the front

# total observations of each
seedling.sums <- colSums(nat.seedlings[7:47], na.rm=TRUE)


# Mia's path
nat.seedlings.10 <- read.csv("C:/Users/mj65tivo/Dropbox/eDiValo-seedlings/Data Entry/natural_regen_2019_10.csv", stringsAsFactors = FALSE)
nat.seedlings.03 <- read.csv("C:/Users/mj65tivo/Dropbox/eDiValo-seedlings/Data Entry/natural_regen_2020_03.csv", stringsAsFactors = FALSE)
nat.seedlings.05 <- read.csv("C:/Users/mj65tivo/Dropbox/eDiValo-seedlings/Data Entry/natural_regen_2020_05.csv", stringsAsFactors = FALSE)
nat.seedlings.06 <- read.csv("C:/Users/mj65tivo/Dropbox/eDiValo-seedlings/Data Entry/natural_regen_2020_06.csv", stringsAsFactors = FALSE)

#View(nat.seedlings)
nat.seedlings[is.na(nat.seedlings)] <- 0

# adding separate columns for exclosure, light, and fertilization treatments
nat.seedlings$grazing <- 'sheep'
nat.seedlings$grazing[grep('E', nat.seedlings$treat)] <- 'exclosure'
nat.seedlings$nutrient <- 'unfertilized'
nat.seedlings$nutrient[grep('F', nat.seedlings$treat)] <- 'fertilized'
nat.seedlings$light <- 'control'
nat.seedlings$light[grep('L', nat.seedlings$treat)] <- 'lamps'
nat.seedlings$block <- as.integer((nat.seedlings$plotid - 1)/8) + 1

# column names other than species
non.species.columns <-c('plotid','treat','block','climate','subplot',
                        'grazing','nutrient','light',
                         'month', 'date') 

# all species column names
species.columns <- names(nat.seedlings)[!(names(nat.seedlings) %in% non.species.columns)] 

# long format of data
nat.seedlings.long <- nat.seedlings %>% 
  pivot_longer(cols = species.columns, names_to = 'species', values_to = 'count')
#View(nat.seedlings.long)

# summary of total seedling count and species richness in each plot (combining subplots)
# split by month
nat.seedlings.plot.sum <- nat.seedlings.long %>% 
  ddply(c('plotid','treat','block','climate','grazing','nutrient','light','month'), 
        summarise,
        total = sum(count),
        richness = sum(count > 0)
  )
nat.seedlings.plot.sum$block <- as_factor(nat.seedlings.plot.sum$block)
nat.seedlings.plot.sum$plotid <- as_factor(nat.seedlings.plot.sum$plotid)
nat.seedlings.plot.sum$month <- as_factor(nat.seedlings.plot.sum$month)

nat.seedlings.plot.sum.2 <- filter(nat.seedlings.plot.sum, 
                                   month %in% c('2019_10','2020_03','2020_05'))
```

## Total seedlings count
Summary for plotting total and richness (code not printed)
```{r summary-plot, echo=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding

# summarize by month and treatments for plotting
nat.seedlings.summary <- nat.seedlings.plot.sum.2 %>% 
  ddply(c('grazing','nutrient','light','month'),
        summarise, 
        n = length(total),
        total.mean = mean(total),
        total.se = sd(total)/sqrt(n),
        rich.mean = mean(richness),
        rich.se = sd(richness)/sqrt(n))

```

Summary for analyzing total and richness (code not printed)
```{r summary-analysis, echo=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding
# splitting data into just fall or just spring for modeling
nat.seedlings.fall <- nat.seedlings.plot.sum %>% filter(month == '2019_10')
nat.seedlings.spring <- nat.seedlings.plot.sum %>% filter(month == '2020_03')
nat.seedlings.late <- nat.seedlings.plot.sum %>% filter(month == '2020_05')

# combined fall and spring together
nat.seedlings.long.fallspring <- nat.seedlings.long %>%
  filter(month %in% c('2019_10', '2020_03'))
nat.seedlings.plot.both <- nat.seedlings.long.fallspring %>% 
  ddply(c('plotid','treat','block','climate','grazing','nutrient','light'), 
        summarise,
        total = sum(count),
        richness = sum(count > 0)
  )
nat.seedlings.plot.both$block <- as_factor(nat.seedlings.plot.both$block)
nat.seedlings.plot.both$plotid <- as_factor(nat.seedlings.plot.both$plotid)
```


### Total seedling count plot
```{r summary-plot-total, echo=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding}
ggplot(nat.seedlings.summary, aes(x=nutrient, y=total.mean, 
                                  ymin=total.mean-total.se,
                                  ymax=total.mean+total.se,
                                  color=light)) +
  facet_grid(month~grazing) +
  geom_point(size=2, position=position_dodge(0.2)) +
  geom_errorbar(width=0.2, position=position_dodge(0.2)) +
  ylab('Total seedling count') +
  xlab('') +
  theme_cw() +
  theme(text = element_text(size=12)) +
  scale_color_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'bottom', legend.text=element_text(size=12))

#ggsave(filename = 'nat_regen_total_time.pdf', width=6, height=5, units='in')
```

### Total seedling analysis

#### With time
```{r total-seedling-time, echo=FALSE, warning=FALSE, message=FALSE}
model.total <- nat.seedlings.plot.sum %>%
  lmer(total ~ grazing + nutrient + light + month +
         grazing:nutrient + grazing:light + grazing:month +
         nutrient:light + nutrient:month + light:month +
         grazing:nutrient:light + 
         (1|block:plotid),.)
tbl_regression(model.total, 
               show_single_row = c('grazing','nutrient','light','grazing:nutrient',
  'grazing:light','nutrient:light','grazing:nutrient:light'))
```


Looking at October and March time points as 1. separate models or 2. a combined sum. I think we can make the argument that if we do it this way we won't need time series analyses since we expect Oct seedlings to have died or grown by March, mostly not resampling the same individuals. But there are other approaches to this that use the time series angle.

#### Fall only
```{r tot-seedling-fall, echo=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding

# want to use block:plot for random effect, but have the error
# number of levels of each grouping factor must be < number of observations
# would need a meta-plot that is the grazing x fert treatment instead
model.total.fall <- nat.seedlings.fall %>%
  lmer(total ~ grazing * nutrient * light + (1|block),.)
tbl_regression(model.total.fall, 
               show_single_row = c('grazing','nutrient','light','grazing:nutrient',
  'grazing:light','nutrient:light','grazing:nutrient:light'))
```

#### Spring only
```{r tot-seedling-spring, echo=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding

model.total.spring <- nat.seedlings.spring %>%
  lmer(total ~ grazing * nutrient * light + (1|block),.)
tbl_regression(model.total.spring, 
               show_single_row = c('grazing','nutrient','light','grazing:nutrient',
                                   'grazing:light','nutrient:light','grazing:nutrient:light'))
```

#### Both combined
```{r tot-seedling-both, echo=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding

model.total.both <- nat.seedlings.plot.both %>%
  lmer(total ~ grazing * nutrient * light + (1|block),.)
tbl_regression(model.total.both, 
               show_single_row = c('grazing','nutrient','light','grazing:nutrient',
                                   'grazing:light','nutrient:light','grazing:nutrient:light'))

```


## Seedling richness
### Seedling richness plot
```{r summary-plot-richness, echo=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding}
ggplot(nat.seedlings.summary, aes(x=nutrient, y=rich.mean, 
                                  ymin=rich.mean-rich.se, 
                                  ymax=rich.mean+rich.se,
                                  color=light)) +
  facet_grid(month~grazing) +
  geom_point(size=2, position=position_dodge(0.2)) +
  geom_errorbar(width=0.2, position=position_dodge(0.2)) +
  ylab('Seedling richness') +
  xlab('') +
  theme_cw() +
  theme(text = element_text(size=12)) +
  scale_color_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'bottom', legend.text=element_text(size=12))
#ggsave(filename = 'nat_regen_richness_time.pdf', width=6, height=5, units='in')
```

### Seedling richness analysis
#### With time
```{r rich-seedling-time, echo=FALSE, warning=FALSE, message=FALSE}
model.richness <- nat.seedlings.plot.sum %>%
  lmer(richness ~ grazing + nutrient + light + month +
         grazing:nutrient + grazing:light + grazing:month +
         nutrient:light + nutrient:month + light:month +
         grazing:nutrient:light + 
         (1|block:plotid),.)
tbl_regression(model.richness, 
               show_single_row = c('grazing','nutrient','light','grazing:nutrient',
  'grazing:light','nutrient:light','grazing:nutrient:light'))
```

#### Fall only
```{r rich-seedling-fall, echo=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding

# want to use block:plot for random effect, but have the error
# number of levels of each grouping factor must be < number of observations
# would need a meta-plot that is the grazing x fert treatment instead
model.richness.fall <- nat.seedlings.fall %>%
  glmer(richness ~ grazing * nutrient * light + (1|block),.,
        family = poisson(link="log"))
tbl_regression(model.richness.fall, 
               show_single_row = c('grazing','nutrient','light','grazing:nutrient',
                                   'grazing:light','nutrient:light','grazing:nutrient:light'))
```

#### Spring only
```{r rich-seedling-spring, echo=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding

model.richness.spring <- nat.seedlings.spring %>%
  glmer(richness ~ grazing * nutrient * light + (1|block),.,
        family = poisson(link="log"))
tbl_regression(model.richness.spring, 
               show_single_row = c('grazing','nutrient','light','grazing:nutrient',
                                   'grazing:light','nutrient:light','grazing:nutrient:light'))
```

#### Both combined
```{r rich-seedling-both, echo=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding

model.richness.both <- nat.seedlings.plot.both %>%
  glmer(richness ~ grazing * nutrient * light + (1|block),.,
        family = poisson(link="log"))
tbl_regression(model.richness.both, 
               show_single_row = c('grazing','nutrient','light','grazing:nutrient',
                                   'grazing:light','nutrient:light','grazing:nutrient:light'))

```

#### April/May 
```{r rich-seedling-late, echo=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding

model.richness.late <- nat.seedlings.late %>%
  glmer(richness ~ grazing * nutrient * light + (1|block),.,
        family = poisson(link="log"))
tbl_regression(model.richness.late, 
               show_single_row = c('grazing','nutrient','light','grazing:nutrient',
                                   'grazing:light','nutrient:light','grazing:nutrient:light'))
```


