---
title: "eDiValo Seedlings Natural Regeneration"
output:
  word_document: default
  html_document: default
---

## Setup and data structuring
Loads libraries, sets up custom theme for ggplot, and reads in data (code not printed)
```{r setup2, include=FALSE, warning=FALSE, message=FALSE}
#NOTE: currently doesn't print warnings or messages, take this out while coding

#library(plyr) # data structuring
library(tidyverse) # data structuring
library(ggplot2) # plotting
library(ggthemes) # plotting
library(lme4) # basic models
library(lmerTest) # p-values on basic models
library(gtsummary) # pretty tables
library(glmmTMB) # glmm with zero-inflation
library(bbmle) # model comparison tables
library(broom.mixed)
source('edivalo-seedlings-traits.R', local=TRUE) # script with the trait data
source('edivalo-seedlings-environmental.R', local = TRUE) # script with env data

# theme for plotting
theme_cw <- function () { 
  theme_bw(base_size=12) %+replace% 
    theme(
      panel.background = element_blank(), 
      plot.background = element_blank(), 
      axis.ticks = element_line(colour = "grey70", size = rel(0.5)),
      panel.grid.minor = element_blank(), 
      panel.grid.major.x = element_blank(),
      legend.background = element_blank(), 
      legend.key = element_blank(),
      strip.background = element_blank(), 
      strip.text=element_text(size=12),
      axis.text=element_text(size=12),
      complete = TRUE
    )
}

nat.seedlings.10 <- read.csv("data/natural_regen_2019_10.csv", 
                             stringsAsFactors = FALSE)
nat.seedlings.03 <- read.csv("data/natural_regen_2020_03.csv", 
                             stringsAsFactors = FALSE)
nat.seedlings.05 <- read.csv("data/natural_regen_2020_05.csv", 
                             stringsAsFactors = FALSE)
nat.seedlings.06 <- read.csv("data/natural_regen_2020_06.csv", 
                             stringsAsFactors = FALSE)


# Want to combine these together
# remove . and spaces
names(nat.seedlings.10) <- gsub("[^a-z]","", names(nat.seedlings.10))
names(nat.seedlings.03) <- gsub("[^a-z]","", names(nat.seedlings.03)) 
names(nat.seedlings.10) <- gsub("[^a-z]","", names(nat.seedlings.10))
names(nat.seedlings.03) <- gsub("[^a-z]","", names(nat.seedlings.03)) 
names(nat.seedlings.05) <- gsub("[^a-z]","", names(nat.seedlings.05))
names(nat.seedlings.06) <- gsub("[^a-z]","", names(nat.seedlings.06))

# small cleaning issues
nat.seedlings.10$subplot <- ifelse(nat.seedlings.10$subplot==1, "O", "I")
nat.seedlings.03$falvul <- as.integer(substr(nat.seedlings.03$falvul, 1, 1))
nat.seedlings.05 <- nat.seedlings.05[,1:43] # removing blank columsn at the end
colnames(nat.seedlings.03)[21]<-"lamsp" # change larnsp to lamsp (likely a typo)
colnames(nat.seedlings.05)[22]<-"lamsp"
colnames(nat.seedlings.06)[22]<-"lamsp"


# add in year_month time point
nat.seedlings.10$month <- '2019_10'
nat.seedlings.03$month <- '2020_03'
nat.seedlings.05$month <- '2020_05'
nat.seedlings.06$month <- '2020_06'

nat.seedlings.a <- full_join(nat.seedlings.10, nat.seedlings.03)
nat.seedlings.b <- full_join(nat.seedlings.05, nat.seedlings.06)
nat.seedlings <- full_join(nat.seedlings.a, nat.seedlings.b)

# move month to the front
nat.seedlings <- nat.seedlings %>% 
  select(month, everything())

# total observations of each
seedling.sums <- colSums(nat.seedlings[7:44], na.rm=TRUE)


#View(nat.seedlings)
nat.seedlings[is.na(nat.seedlings)] <- 0

# adding separate columns for exclosure, light, and fertilization treatments
nat.seedlings$grazing <- 'sheep'
nat.seedlings$grazing[grep('E', nat.seedlings$treat)] <- 'exclosure'
nat.seedlings$nutrient <- 'unfertilized'
nat.seedlings$nutrient[grep('F', nat.seedlings$treat)] <- 'fertilized'
nat.seedlings$light <- 'control'
nat.seedlings$light[grep('L', nat.seedlings$treat)] <- 'lamps'
nat.seedlings$block <- as.integer((nat.seedlings$plotid - 1)/8) + 1

# column names other than species
non.species.columns <-c('plotid','treat','block','climate','subplot',
                        'grazing','nutrient','light',
                         'month', 'date') 

# all species column names
species.columns <- names(nat.seedlings)[!(names(nat.seedlings) %in% non.species.columns)] 

# long format of data
# also grouping the subplots together here
nat.seedlings.long <- nat.seedlings %>% 
  pivot_longer(cols = species.columns, 
               names_to = 'species', values_to = 'count') %>%
  group_by(plotid, treat, block, climate, grazing, nutrient, light, species, month) %>%
  dplyr::summarise(
    count = sum(count)
  )
#View(nat.seedlings.long)

# summary of total seedling count and species richness in each plot
# split by month
nat.seedlings.plot.sum <- nat.seedlings.long %>% 
  group_by(plotid, treat, block, climate, grazing, nutrient, light, month) %>%
  dplyr::summarise(
    total = sum(count),
    richness = sum(count > 0)
  )

nat.seedlings.plot.sum$block <- as_factor(nat.seedlings.plot.sum$block)
nat.seedlings.plot.sum$plotid <- as_factor(nat.seedlings.plot.sum$plotid)
nat.seedlings.plot.sum$month <- as_factor(nat.seedlings.plot.sum$month)

nat.seedlings.plot.sum.2 <- filter(nat.seedlings.plot.sum, 
                                   month %in% c('2019_10','2020_03','2020_05'))
```


## Seedling Traits

For the trait analyses we will look at the max number of seedlings per species across the time points (peak emergence for each species)
```{r trait-biomass, echo=FALSE, warning=FALSE, message=FALSE}

# max across time points
nat.seedlings.max <- nat.seedlings.long %>%
  group_by(plotid, treat, block, grazing, nutrient, light, species, climate) %>%
  dplyr::summarise(
    count.max = max(count)
  )

# combine with trait data, 
# filter down to the species for which we have trait data
nat.seedlings.traits <- nat.seedlings.max %>% 
  left_join(species.size.7, by = 'species') %>% 
  filter(!is.na(bm.tot))
nat.seedlings.traits$block <- as.factor(nat.seedlings.traits$block)
nat.seedlings.traits$plotid <- as.factor(nat.seedlings.traits$plotid)

# simulated dataset for counter-factual plots
new.data <- nat.seedlings.traits
new.data$species <- "new"
new.data$block <- "new"
new.data$plotid <- "new"


# printing species and trait dataset
nat.seedlings.traits.b <- nat.seedlings.max %>% 
  full_join(species.size.7, by = 'species')

nat.seedlings.traits.sum.3 <- nat.seedlings.traits.b %>%
  group_by(species) %>%
  dplyr::summarise(
    abundance = mean(count.max),
    abundance.se = sd(count.max)/sqrt(20),
    bm.tot = mean(bm.tot), 
    rt.sh.bm = mean(rt.sh.bm),
    len.sh = mean(len.sh),
    sla = mean(sla.2),
    c.n = mean(c.n.ratio)
    # traits should be the same as their values, double-check
  )

nat.seedlings.traits.sum.3 %>% 
  arrange(-abundance, species) %>%
  write.csv(file = 'species_traits.csv')

```

### Species traits abundance models
The data is overdispersed and/or zero-inflated. Using the 'glmmTMB' package, which is detailed in "glmmTMB Balances Speed and Flexibility Among Packages for Zero-inflated Generalized Linear Mixed Modeling" by Brooks et al 2017, and the AICtab() function in the 'bblme' package for convenient AIC comparisons of different models. 

Running analyses for now with each trait separately. Could also consider models that include multiple traits, especially since most of the traits aren't strongly correlated with each other. 

For each trait, we are comparing five models:
0 -- null model without the traits, with a random variable for species
1 -- just the trait, no random variable for species 
2 -- trait and random variable for species
full -- all 2-way, 3-way, and 4-way interactions
specific -- 3-way interactions for grazing x light x trait and grazing x light x nutrients, which are the ones that are significant in some of the models. 

Results below each model show the summary for the specific model, an AIC table comparing these options, and the under- or over-dispersion coefficient. 

### Total biomass 

Analysis of total biomass trait using a negative binomial distribution. We do see some (although not particularly strong patterns). Tested to see if this was robust to removing lamium from the dataset (which has a MUCH higher biomass than everything else) and yes, the values are nearly identical with and without it. 
```{r trait-biomass-analysis, echo=FALSE, warning=FALSE, message=FALSE}
# TEST: removing lamium to see if the patterns hold up
#nat.seedlings.traits.no.lam <- filter(nat.seedlings.traits, !(species == 'lamsp'))


m.biomass.clim.0 <- glmmTMB(count.max ~ grazing + nutrient + climate + light +
                         (1|species) + (1|block:plotid),
                       nat.seedlings.traits, family = nbinom2)

m.biomass.clim.1 <- glmmTMB(count.max ~ grazing + nutrient + climate + light + bm.tot +
                         (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)


m.biomass.clim.full <- glmmTMB(count.max ~ grazing + nutrient + climate + light + bm.tot +
                           grazing*nutrient*climate*light*bm.tot +
                         (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

# summary(m.biomass.clim.full)

# remove non-significant interactions
m.biomass.clim.specific <- glmmTMB(count.max ~ grazing + nutrient + 
                                     climate + light + bm.tot +
                           nutrient*grazing*bm.tot +
                         (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)
summary(m.biomass.clim.specific)

E2 <- resid(m.biomass.clim.specific, type = "pearson")
N  <- nrow(nat.seedlings.traits)
p  <- length(coef(m.biomass.clim.specific)) 
sum(E2^2) / (N - p)

# including climate-trait effect
# remove non-significant interactions
m.biomass.clim.specific.2 <- glmmTMB(count.max ~ grazing + nutrient + 
                                     climate + light + bm.tot +
                           nutrient*grazing*bm.tot + climate*bm.tot + 
                         (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)
summary(m.biomass.clim.specific.2)

AICtab(m.biomass.clim.0, m.biomass.clim.1, m.biomass.clim.full, m.biomass.clim.specific, m.biomass.clim.specific.2)

```

Counterfactual plot of biomass x grazing x nutrient interaction. Using re.form = NA to specific population-level predictions (setting all random effects to zero)
```{r trait-biomass-plots2, echo=FALSE, warning=FALSE, message=FALSE}
# TEST: removing lamium to see if the patterns hold up
#nat.seedlings.traits.no.lam <- filter(nat.seedlings.traits, !(species == 'lamsp'))
#new.data.no.lam <- filter(new.data, bm.tot < 100)

# simplified (without light or climate) for plotting
m.biomass.simple <- glmmTMB(count.max ~ nutrient*grazing*bm.tot + 
                          (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

m.biomass.fit <- predict(m.biomass.simple, new.data, allow.new.levels = TRUE, se.fit = TRUE,
                     re.form = NA, type = 'response')
new.data$biomass.fit <- m.biomass.fit$fit
new.data$biomass.fit.se <- m.biomass.fit$se.fit

# Easier to visualize the other way around 
# In fertilized plots, trait relationship doesn't depend on grazing
# but in unfertilized plots, it does (sharp threshold in exclosure plots)

ggplot(NULL) + 
  facet_grid(.~nutrient) +
  geom_line(data = new.data, size = 1.2, 
            aes(x = bm.tot, y = biomass.fit, color = grazing)) + 
  geom_ribbon(data = new.data, alpha = 0.3,
              aes(x = bm.tot, y = biomass.fit, 
                  ymin = biomass.fit - biomass.fit.se, ymax = biomass.fit + biomass.fit.se,
                  fill = grazing)) +
  # geom_point(data = nat.seedlings.traits.sum.2, aes(x = bm.tot, y = abundance, color = light)) + 
  theme_cw() + 
  theme(text = element_text(size=12)) +
  #  scale_color_manual(values=c('black','orange'), name='') +
  #  scale_fill_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'bottom', legend.text=element_text(size=12)) + 
  ylab('Predicted abundance') + 
  xlab('Biomass')

# one more version, just looking at the nutrient x biomass interaction
m.biomass.simple.2 <- glmmTMB(count.max ~ nutrient*bm.tot + 
                          (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

m.biomass.fit.2 <- predict(m.biomass.simple.2, new.data, allow.new.levels = TRUE, se.fit = TRUE,
                     re.form = NA, type = 'response')
new.data$biomass.fit.2 <- m.biomass.fit.2$fit
new.data$biomass.fit.se.2 <- m.biomass.fit.2$se.fit

ggplot(NULL) + 
  geom_line(data = new.data, size = 1.2, 
            aes(x = bm.tot, y = biomass.fit.2, color = nutrient)) + 
  geom_ribbon(data = new.data, alpha = 0.3,
              aes(x = bm.tot, y = biomass.fit.2, 
                  ymin = biomass.fit.2 - biomass.fit.se.2, 
                  ymax = biomass.fit.2 + biomass.fit.se.2,
                  fill = nutrient)) +
  # geom_point(data = nat.seedlings.traits.sum.2, aes(x = bm.tot, y = abundance, color = light)) + 
  theme_cw() + 
  theme(text = element_text(size=12)) +
  #  scale_color_manual(values=c('black','orange'), name='') +
  #  scale_fill_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'bottom', legend.text=element_text(size=12)) + 
  ylab('Predicted abundance') + 
  xlab('Biomass')

```

### SLA 

Analysis of SLA
```{r trait-sla-analysis, echo=FALSE, warning=FALSE, message=FALSE}

# including climate
m.sla.clim.0 <- glmmTMB(count.max ~ grazing + nutrient + light + climate +
                         (1|species) + (1|block:plotid),
                       nat.seedlings.traits, family = nbinom2)


m.sla.clim.1 <- glmmTMB(count.max ~ grazing + nutrient + light + climate + sla.2 +
                         (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

m.sla.clim.full <- glmmTMB(count.max ~ grazing + nutrient + light + climate + sla.2 +
                            grazing*nutrient*light*climate*sla.2 +
                        (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)
# summary(m.sla.clim.full)

m.sla.clim.specific <- glmmTMB(count.max ~ grazing + nutrient + light + climate + sla.2 +
                           grazing*nutrient*light + grazing*nutrient*climate +
                             grazing*nutrient*sla.2 + grazing*light*sla.2 +
                             nutrient*light*sla.2 + nutrient*climate*sla.2 +
                             grazing*nutrient*climate*sla.2 +
                        (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)
summary(m.sla.clim.specific) # yikes, 4-way interactions. Hard to visualize/interpret



AICtab(m.sla.clim.0, m.sla.clim.1, m.sla.clim.full, m.sla.clim.specific)

E2 <- resid(m.sla.clim.specific, type = "pearson")
N  <- nrow(nat.seedlings.traits)
p  <- length(coef(m.sla.clim.specific)) 
sum(E2^2) / (N - p) 

```

Counterfactual plots Using re.form = NA to specific population-level predictions (setting all random effects to zero)
```{r trait-SLA-plots2, echo=FALSE, warning=FALSE, message=FALSE}


# plot 1: looking at the 4-way grazing-nutrient-climate-sla interaction
m.sla.plot1<- glmmTMB(count.max ~ grazing*nutrient*climate*sla.2+ 
                          (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

m.sla.fit1 <- predict(m.sla.plot1, new.data, allow.new.levels = TRUE, se.fit = TRUE,
                     re.form = NA, type = 'response')
new.data$sla.fit1 <- m.sla.fit1$fit
new.data$sla.fit1.se <- m.sla.fit1$se.fit

ggplot(NULL) + 
  facet_grid(climate~nutrient) +
  geom_line(data = new.data, size = 1.2, aes(x = sla.2, y = sla.fit1, color = grazing)) + 
  geom_ribbon(data = new.data, alpha = 0.3,
              aes(x = sla.2, y = sla.fit1, 
                  ymin = sla.fit1 - sla.fit1.se, ymax = sla.fit1 + sla.fit1.se,
                  fill = grazing)) +
#  geom_point(data = nat.seedlings.traits.sum.2, aes(x = sla, y = abundance, color = light)) + 
  theme_cw() + 
  theme(text = element_text(size=12)) +
#  scale_color_manual(values=c('black','orange'), name='') +
#  scale_fill_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'bottom', legend.text=element_text(size=12)) + 
  ylab('Predicted abundance') + 
  xlab('SLA')
#ggsave('sla-model-data.pdf', width = 6, height = 4, units = 'in')



# plot 2: looking at the 3-way grazing-light-sla interaction
m.sla.plot2<- glmmTMB(count.max ~ light*grazing*sla.2 + 
                          (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

m.sla.fit2 <- predict(m.sla.plot2, new.data, allow.new.levels = TRUE, se.fit = TRUE,
                     re.form = NA, type = 'response')
new.data$sla.fit2 <- m.sla.fit2$fit
new.data$sla.fit2.se <- m.sla.fit2$se.fit


ggplot(NULL) + 
  facet_grid(.~grazing) +
  geom_line(data = new.data, size = 1.2, aes(x = sla.2, y = sla.fit2, color = light)) + 
  geom_ribbon(data = new.data, alpha = 0.3,
              aes(x = sla.2, y = sla.fit2, 
                  ymin = sla.fit2 - sla.fit2.se, ymax = sla.fit2 + sla.fit2.se,
                  fill = light)) +
#  geom_point(data = nat.seedlings.traits.sum.2, aes(x = sla, y = abundance, color = light)) + 
  theme_cw() + 
  theme(text = element_text(size=12)) +
  scale_color_manual(values=c('black','orange'), name='') +
  scale_fill_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'bottom', legend.text=element_text(size=12)) + 
  ylab('Predicted abundance') + 
  xlab('SLA')
#ggsave('sla-model-data.pdf', width = 6, height = 4, units = 'in')


```



### C:N 

```{r trait-cn-analysis, echo=FALSE, warning=FALSE, message=FALSE}
m.cn.clim.0 <- glmmTMB(count.max ~ grazing + nutrient + light + climate +
                         (1|species) + (1|block:plotid),
                       nat.seedlings.traits, family = nbinom2)


m.cn.clim.1 <- glmmTMB(count.max ~ grazing + nutrient + light + climate + c.n.ratio +
                         (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

m.cn.clim.full <- glmmTMB(count.max ~ grazing + nutrient + light + climate + c.n.ratio +
                            grazing*nutrient*light*climate*c.n.ratio +
                        (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)
# summary(m.cn.clim.full)

m.cn.clim.specific <- glmmTMB(count.max ~ grazing + nutrient + light + climate + c.n.ratio +
                            grazing*light*climate*c.n.ratio + nutrient*c.n.ratio +
                        (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

summary(m.cn.clim.specific)

AICtab(m.cn.clim.0, m.cn.clim.1, m.cn.clim.full, m.cn.clim.specific)


```

Counterfactual plots
```{r trait-CN-plots2, echo=FALSE, warning=FALSE, message=FALSE}

cn.data <- new.data[complete.cases(new.data),] 

m.cn.plot1 <- glmmTMB(count.max ~ light*grazing*climate*c.n.ratio + 
                          (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

m.cn.fit1 <- predict(m.cn.plot1, cn.data, allow.new.levels = TRUE, se.fit = TRUE,
                     re.form = NA, type = 'response')
cn.data$cn.fit1 <- m.cn.fit1$fit
cn.data$cn.fit1.se <- m.cn.fit1$se.fit

ggplot(NULL) + 
  facet_grid(climate~light) +
  geom_line(data = cn.data, size = 1.2, aes(x = c.n.ratio, y = cn.fit1, color = grazing)) + 
  geom_ribbon(data = cn.data, alpha = 0.3,
              aes(x = c.n.ratio, y = cn.fit1, 
                  ymin = cn.fit1 - cn.fit1.se, ymax = cn.fit1 + cn.fit1.se,
                  fill = grazing)) +
#  geom_point(data = nat.seedlings.traits.sum.2, aes(x = c.n, y = abundance, color = light)) + 
  theme_cw() + 
  theme(text = element_text(size=12)) +
#  scale_color_manual(values=c('black','orange'), name='') +
#  scale_fill_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'bottom', legend.text=element_text(size=12)) + 
  ylab('Predicted abundance') + 
  xlab('C:N')
#ggsave('cn-model.pdf', width = 6, height = 4, units = 'in')

m.cn.plot2 <- glmmTMB(count.max ~ nutrient*c.n.ratio + 
                          (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

m.cn.fit2 <- predict(m.cn.plot2, cn.data, allow.new.levels = TRUE, se.fit = TRUE,
                     re.form = NA, type = 'response')
cn.data$cn.fit2 <- m.cn.fit2$fit
cn.data$cn.fit2.se <- m.cn.fit2$se.fit

ggplot(NULL) + 
#  facet_grid(climate~light) +
  geom_line(data = cn.data, size = 1.2, aes(x = c.n.ratio, y = cn.fit2, color = nutrient)) + 
  geom_ribbon(data = cn.data, alpha = 0.3,
              aes(x = c.n.ratio, y = cn.fit2, 
                  ymin = cn.fit2 - cn.fit2.se, ymax = cn.fit2 + cn.fit2.se,
                  fill = nutrient)) +
#  geom_point(data = nat.seedlings.traits.sum.2, aes(x = c.n, y = abundance, color = light)) + 
  theme_cw() + 
  theme(text = element_text(size=12)) +
#  scale_color_manual(values=c('black','orange'), name='') +
#  scale_fill_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'bottom', legend.text=element_text(size=12)) + 
  ylab('Predicted abundance') + 
  xlab('C:N')

```

### Root-shoot ratio
no specific model, since nothing is significant (not even a general relationship with the trait)
```{r trait-rtsh-analysis, echo=FALSE, warning=FALSE, message=FALSE}

m.rt.sh.bm.0 <- glmmTMB(count.max ~ grazing * nutrient + climate + light +
                         (1|species) + (1|block:plotid),
                       nat.seedlings.traits, family = nbinom2)

m.rt.sh.bm.1 <- glmmTMB(count.max ~ grazing * nutrient + climate + light + rt.sh.bm +
                         (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

m.rt.sh.bm.full <- glmmTMB(count.max ~ grazing * nutrient + climate + light + rt.sh.bm +
                            grazing*nutrient*climate*light*rt.sh.bm +
                        (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)


AICtab(m.rt.sh.bm.0, m.rt.sh.bm.1, m.rt.sh.bm.full)

```

### Shoot length
no specific model, since nothing is significant (not even a general relationship with the trait)

```{r trait-len-analysis, echo=FALSE, warning=FALSE, message=FALSE}


m.len.sh.0 <- glmmTMB(count.max ~ grazing * nutrient + climate + light +
                         (1|species) + (1|block:plotid),
                       nat.seedlings.traits, family = nbinom2)

m.len.sh.1 <- glmmTMB(count.max ~ grazing * nutrient + climate + light + len.sh +
                         (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

m.len.sh.full <- glmmTMB(count.max ~ grazing * nutrient + climate +light + len.sh +
                            grazing*nutrient*climate*light*len.sh +
                        (1|species) + (1|block:plotid), 
                 nat.seedlings.traits, family = nbinom2)

# nothing significant for specific 

AICtab(m.len.sh.0, m.len.sh.1, m.len.sh.full)
```





