---
title: "edivalo-seedlings-toothpicks"
output:
  word_document: default
  html_document: default
---

## Setup and data structuring
Loads libraries, sets up custom theme for ggplot, and reads in data (code not printed)
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
#library(plyr)
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(plyr) # hoping we can cut this by cleaning up summary code
library(glmmTMB) # for emergence analysis
library(coxme) # for survival analysis
library(car) # for survival analysis
library(survminer) # for survival plots

# theme for plotting
theme_cw <- function () { 
  theme_bw(base_size=10) %+replace% 
    theme(
      panel.background = element_blank(), 
      plot.background = element_blank(), 
      axis.ticks = element_line(colour = "grey70", size = rel(0.5)),
      panel.grid.minor = element_blank(), 
      panel.grid.major.x = element_blank(),
      legend.background = element_blank(), 
      legend.key = element_blank(),
      strip.background = element_blank(), 
      strip.text=element_text(size=10),
      axis.text=element_text(size=10),
      complete = TRUE
    )
}


mark.seedlings <- read.csv("data/toothpick_seedling_data_2020_all.csv",
                           stringsAsFactors = FALSE)
plot.list <- read.csv("data/plot_list.csv", stringsAsFactors = FALSE)

# Mia's path
#mark.seedlings<-read.csv("C:/Users/mj65tivo/Documents/GitHub/edivalo-seedlings/data/toothpick_seedling_data_2020_all.csv", stringsAsFactors = FALSE)
#View(mark.seedlings)
# For Mia: change first column name to month
#colnames(mark.seedlings)[1]<-"month"



# adding separate columns for exclosure, light, and fertilization treatments
mark.seedlings$grazing <- 'sheep'
mark.seedlings$grazing[grep('E', mark.seedlings$plot)] <- 'exclosure'
mark.seedlings$nutrient <- 'unfertilized'
mark.seedlings$nutrient[grep('F', mark.seedlings$plot)] <- 'fertilized'
mark.seedlings$light <- 'control'
mark.seedlings$light[grep('L', mark.seedlings$plot)] <- 'lamps'

# summary of total number by species, plot, and month
mark.summary <- mark.seedlings %>%
    group_by(block, plot, grazing, nutrient, light, species, month) %>%
  dplyr::summarize(
    count = sum(number),
    count.2 = sum(number>0))

# combine plot numbers and treatments together in plot list file
plot.list$plot <- paste(plot.list$plotid, plot.list$treat, sep="")
plot.list$block <- rep(1:10, each = 8)

```

## Visualization plots

### All species through time
Taking a first look at our data, just the abundance counts of the six species through time. Months are listed on the right side (2-February to 7-July). Seedling abundance was generally highest in March-April-May, and the most common species were Plantago lanceolata and Dianthus. 
```{r full, echo=FALSE, warning=FALSE, message=FALSE}
mark.plot <- mark.summary %>% 
  group_by(grazing, nutrient, light, species, month) %>%
  dplyr::summarize(
    plots.present = length(count),
    total.mean = sum(count)/10,
    total.se = sd(count)/sqrt(10))

plantago <- filter(mark.plot, species=="plantago lan")
daucus <- filter(mark.plot, species=="daucus car")
medicago <- filter(mark.plot, species=="medicago fal.")
dianthus <- filter(mark.plot, species=="dianthus car")
galium <- filter(mark.plot, species=="galium alb")
crepis <- filter(mark.plot, species=="crebis bie")

## everything
ggplot(mark.plot, aes(x=grazing, y=total.mean, 
                     ymin=total.mean-total.se, ymax=total.mean+total.se,
                     color=light, shape=nutrient)) +
  facet_grid(month~species) +
  geom_point(size=2, position=position_dodge(0.2)) +
  geom_errorbar(width=0.2, position=position_dodge(0.2)) +
  ylab('Seedling count') +
  xlab('') +
  theme_cw() +
  theme(text = element_text(size=10)) +
  scale_color_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'top', legend.text=element_text(size=10))

#ggsave(filename = 'species_time.pdf', width=10, height=8, units='in')

```


## Emergence
Data re-structuring to look at emergence rather than abundance. We also want to add in zeros for the ones that are missing for the analyses, we're using a complete plot list for this. It is multiplied to have all the species in each plot. 

```{r emergence-data, echo=FALSE, warning=FALSE, message=FALSE}

# for each toothpick, if one or two seedlings emerged
mark.emergence <- mark.seedlings %>%
  group_by(block, plot, grazing, nutrient, light, species, id..1.6.) %>%
  dplyr::summarize(
    emergence.2 = max(number),
    emergence.1 = ifelse(emergence.2 > 0, 1, 0)
    )

# summary of number of seedlings emerged per species per plot
emergence.plot <- mark.emergence %>%
  group_by(block, plot, grazing, nutrient, light, species) %>%
  dplyr::summarise(
    emerged = sum(emergence.2),
    emerged.1 = sum(emergence.1)
  )

# setting up the full data frame with zeros
species.list <- unique(mark.plot$species)
plot.rep <- rep.int(plot.list$plot, length(species.list))
block.rep <- rep.int(plot.list$block, length(species.list))
species.rep <- rep(species.list, each=length(plot.list$plot))

emergence.plot.min <- emergence.plot %>%
  ungroup() %>%
  select(., plot, species, emerged, emerged.1)

emergence.full <- data.frame(plot = plot.rep, species = species.rep, block = block.rep) %>%
  left_join(emergence.plot.min, by = c('plot', 'species'))

# consider turning this into a function?
emergence.full$grazing <- 'sheep'
emergence.full$grazing[grep('E', emergence.full$plot)] <- 'exclosure'
emergence.full$nutrient <- 'unfertilized'
emergence.full$nutrient[grep('F', emergence.full$plot)] <- 'fertilized'
emergence.full$light <- 'control'
emergence.full$light[grep('L', emergence.full$plot)] <- 'lamps'

# changing added 'na' rows to zeros
emergence.full$emerged[is.na(emergence.full$emerged)] <- 0
emergence.full$emerged.1[is.na(emergence.full$emerged.1)] <- 0

```

### Emergence Plot
Plotting emergence by species and treatment
```{r emergence-plot, echo=FALSE, warning=FALSE, message=FALSE}

emergence.sum  <- emergence.full %>%
  group_by(grazing, nutrient, light, species) %>%
  dplyr::summarize(
    n = length(emerged),
    emergence.mean = mean(emerged, na.rm = TRUE),
    emergence.se = sd(emerged)/sqrt(n)
  )

ggplot(emergence.sum,
       aes(x=grazing, y=emergence.mean, 
           ymin=emergence.mean-emergence.se, ymax=emergence.mean+emergence.se,
           color=light, shape = nutrient)) +
  facet_wrap(vars(species)) +
  geom_point(size=3, position=position_dodge(0.3)) +
  geom_errorbar(width=0.1, position=position_dodge(0.3)) +
  ylab('Seedling emergence') +
  xlab('') +
  theme_cw() +
  theme(text = element_text(size=10)) +
  scale_color_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'right', legend.text=element_text(size=10))

```

### Emergence Analysis
Separate model for each species, only looking at dianthus and plantago. Currently using negative binomial, seem to be somewhat underdispersed

```{r emergence-analysis, echo=FALSE, warning=FALSE, message=FALSE}

emerg.pla <- emergence.full %>% filter(species == 'plantago lan')
emerg.dia <- emergence.full %>% filter(species == 'dianthus car')

m.emergence.pla <- glmmTMB(emerged ~ grazing * light + nutrient*light +
                     (1|block), 
                   emerg.pla, family = nbinom2)
summary(m.emergence.pla)
E2 <- resid(m.emergence.pla, type = "pearson")
N  <- nrow(emerg.pla)
p  <- length(coef(m.emergence.pla)) 
sum(E2^2) / (N - p) 

m.emergence.dia <- glmmTMB(emerged ~ grazing * light + nutrient*light +
                     (1|block), 
                   emerg.dia, family = nbinom2)
summary(m.emergence.dia)
E2 <- resid(m.emergence.dia, type = "pearson")
N  <- nrow(emerg.dia)
p  <- length(coef(m.emergence.dia)) 
sum(E2^2) / (N - p) 

```


## Survival
Survival structuring 

Below is the code Mia had. Some of this wasn't working or didn't make sense to me, so I ended up taking a different approach. I'm keeping this here in case reverse changes need to be made :-) 

Questions: 1. Are the days based on the actual days we sampled? 2. Cleaner way to do the sampling time etc without all the ifelse statements (ex matching to a vector) 3. Right now the "event" column is just the opposite of the "alive" column. I think it's supposed to be the first time a seedling dies, which takes a little more work. But I need to check with this package, I think it's different from the ones I used before (survival + survminer)
```{r survival-unused, echo = FALSE}
# previous structuring work Mia did
# Something is wrong with the replication order and the blocks/plots aren't lining up
# so the join isn't working correctly. 
#To do it the way you have it, you need to check the sorting first, right now
# the plots are getting assigned to the wrong blocks. 

# # setting up the full data frame with one line for every toothpick at each sampling point
# species.list <- unique(mark.seedlings$species) # should be 6
# toothpick.list <-unique(mark.seedlings$id..1.6.) # should be 6
# month.list<-unique(mark.seedlings$month) # should be 6
# plot.list.2 <- unique(plot.list$plot)# should be 80
# block.list <- unique(mark.seedlings$block)
# 
# 
# month.rep1 <- rep(month.list, each= 2880, times = 1)
# block.rep1 <- rep(block.list, each= 288, times= 6)
# plot.rep1 <- rep(plot.list.2, each=36, times = length(month.list))
# species.rep1 <- rep(species.list, each=length(toothpick.list), times = 480)
# toothpick.rep1 <- rep(toothpick.list, times = 2880)
# 
# ## NOTE from Chhaya: mark.alive is exactly the same as mark.seedlings? 
# ## What was the goal of this step? 
# 
# # for each toothpick and sampling time (month) how many seedlings are there (alive)
#  mark.alive <- mark.seedlings %>%
#   group_by(month,block, plot, grazing, nutrient, light, species, id..1.6.) %>%
#   dplyr::summarize(
#     alive = number)
# 
# 
# ## NOTE from Chhaya: we actually only want a data frame for all the seedlings that did emerge
# ## If they didn't emerge they shouldn't be included in the survival analysis
# 
# # create full data frame to also include the toothpicks were no seedlings were alive/emerged
# surv.full <-  data.frame(month = month.rep1, block = block.rep1, plot = plot.rep1, 
#                          species = species.rep1, id..1.6. = toothpick.rep1) %>%
#   left_join(mark.seedlings)


# surv.full$alive[is.na(surv.full$alive)] <- 0
# surv.full$alive = as.numeric(surv.full$alive)
# 
# surv.full$grazing <- 'sheep'
# surv.full$grazing[grep('E', surv.full$plot)] <- 'exclosure'
# surv.full$nutrient <- 'unfertilized'
# surv.full$nutrient[grep('F', surv.full$plot)] <- 'fertilized'
# surv.full$light <- 'control'
# surv.full$light[grep('L', surv.full$plot)] <- 'lamps'

# #  add column for sampling time, day 
# # and the beginning of the interval in which a seedling could have died 
# surv.full$sampling = ifelse(surv.full$month == "2", "t1",
#                           ifelse(surv.full$month == "3", "t2",
#                                  ifelse(surv.full$month == "4", "t3",
#                                         ifelse(surv.full$month == "5", "t4",
#                                                ifelse(surv.full$month == "6", "t5", "t6")))))
# 
# surv.full$doy = ifelse(surv.full$month == "2", "50",
#                           ifelse(surv.full$month == "3", "77",
#                                  ifelse(surv.full$month == "4", "107",
#                                         ifelse(surv.full$month == "5", "147",
#                                                ifelse(surv.full$month == "6", "171", "198")))))
# 
# surv.full$int_start = ifelse(surv.full$doy == "50", "0",
#                                    ifelse(surv.full$month == "77", "50",
#                                       ifelse(surv.full$month == "107", "77",
#                                         ifelse(surv.full$month == "147", "107",
#                                               ifelse(surv.full$month == "171","147", "171")))))
# 
# # remove those seedlings that are always zero through month 2,3,4,6,7
# # THIS STILL NEEDS TO BE DONE
# surv.full1 <- surv.full %>% 
#   filter(!(month =="2" & alive == "0"))
# 
# # add row with with event (seedling dead = 1, still alive = 0)
# surv.full1$event <- ifelse(surv.full1$alive == "0", 1, 0)

## the above keeps 0s for both seedlings that have emerged and died 
## and seedlings that haven't emerged yet--is this how it should be?

```

I think we want to limit this to the toothpicks where seedlings did emerge, not all of them. So I'm going to start with the mark.emergence data set, and then increase it to have the data needed for each month. We can use the tidyr function crossing() to expand it to the size we need, and then fill in the number column by merging with the mark.seedlings dataset and setting the NA's to 0 (when a seedling wasn't observed). 


```{r survival-analysis}

surv.full <- data.frame(month = 2:7) %>% 
  crossing(mark.emergence) %>% 
  left_join(mark.seedlings) %>%
  replace_na(list(number = 0))

surv.full$alive <- ifelse(surv.full$number > 0, 1, 0)


# this doesn't use the full time intervals the way Mia has it set up
surv.full.2 <- surv.full %>%
  group_by(block, plot, grazing, nutrient, light, species, id..1.6.) %>%
  dplyr::summarise(
    dead = 1 - alive[month == max(month)], # at the end dead (1) or alive (0)
    month.emerged = min(month[alive > 0]),
    month.died = max(month[alive > 0]),
    months.alive = (month.died - month.emerged) + 1, # not sure if we want the +1 or not
    months.alive.2 = sum(alive)
  )

surv.pla <- filter(surv.full.2, species == 'plantago lan')
surv.dia <- filter(surv.full.2, species == 'dianthus car')

surv.obj.pla <- Surv(surv.pla$months.alive, surv.pla$dead, type = 'right')
surv.obj.dia <- Surv(surv.dia$months.alive, surv.dia$dead, type = 'right')

```

Running models for the survival using the coxme package. 

Here I used right-censored models (number of months alive, whether or not they're currently dead). Mia uses a different survival structure, with day of year (could be month or day, that part isn't super important) and event of dying, but this doesn't include information on when they germinated?

The Anova part doesn't seem to be working? But summary() has p-values so I'm not sure why we need it? 
```{r survival-models}

# Mia: We might need to add another level of hierachy in the random term?

# Dianthus--doesn't run if 3-way interaction is used
m.surv.dia <- coxme(surv.obj.dia ~ grazing*light + nutrient*light + (1|block/plot), 
                  data = surv.dia)
summary(m.surv.dia) # nothing sigificant
#Anova(m.surv.dia, type = "II", test.statistic = "Chisq")

# Plantago
m.surv.pla <- coxme(surv.obj.pla ~ grazing*light + nutrient*light + (1|block/plot), 
                  data = surv.pla)
summary(m.surv.pla) # nothing significant
# Anova(m.surv.pla, type = "II", test.statistic = "Chisq") this doesn't run?

```

### Survival plots
Trying to take a look at the survival patterns, just for plantago. Plotting package doesn't take random effects (ex block) into account
```{r survival-plots-pla}

# this just makes the legend of the plot easier to see
surv.pla.2 <- surv.pla
surv.pla.2$sheep <- ifelse(surv.pla$grazing=='sheep', 1, 0)
surv.pla.2$fert <- ifelse(surv.pla$nutrient=='fertilized', 1, 0)
surv.pla.2$lamp <- ifelse(surv.pla$light=='lamps', 1, 0)

surv.fit.pla <- survfit(surv.obj.pla ~ sheep + fert + lamp, 
                  data = surv.pla.2)
ggsurvplot(surv.fit.pla, 
           palette=c('black','orange','black','orange',
                     'black','orange','black','orange'), 
           linetype = c('longdash', 'longdash','solid', 'solid', 
                        'dotted','dotted', 'dashed', 'dashed'), 
           ggtheme = theme_cw(), 
           xlab='Months Since Emergence')


# just grazing and sheep
surv.fit.pla.2 <- survfit(surv.obj.pla ~ sheep + lamp, 
                  data = surv.pla.2)
ggsurvplot(surv.fit.pla.2, 
           palette=c('black','orange','black','orange'), 
           linetype = c('solid', 'solid', 'dashed', 'dashed'), 
           ggtheme = theme_cw(), 
           xlab='Months Since Emergence')

```


Same for dianthus
```{r survival-plots-dia}

# this just makes the legend of the plot easier to see
surv.dia.2 <- surv.dia
surv.dia.2$sheep <- ifelse(surv.dia$grazing=='sheep', 1, 0)
surv.dia.2$fert <- ifelse(surv.dia$nutrient=='fertilized', 1, 0)
surv.dia.2$lamp <- ifelse(surv.dia$light=='lamps', 1, 0)

surv.fit.dia <- survfit(surv.obj.dia ~ sheep + fert + lamp, 
                  data = surv.dia.2)

## apparently there were no dianthus seedlings in one of the conditions
## (no sheep, fertilized, lamps)
ggsurvplot(surv.fit.dia, 
           palette=c('black','orange','black',
                     'black','orange','black','orange'), 
           linetype = c('longdash', 'longdash','solid', 
                        'dotted','dotted', 'dashed', 'dashed'), 
           ggtheme = theme_cw(), 
           xlab='Months Since Emergence')


# just grazing and sheep
# this makes it look like lamps may have had an effect
# I think looking at a continuous spectrum of light could be a good option moving forward>
surv.fit.dia.2 <- survfit(surv.obj.dia ~ sheep + lamp, 
                  data = surv.dia.2)
ggsurvplot(surv.fit.dia.2, 
           palette=c('black','orange','black','orange'), 
           linetype = c('solid', 'solid', 'dashed', 'dashed'), 
           ggtheme = theme_cw(), 
           xlab='Months Since Emergence')

```

## Next steps
Might be worth taking another look at the survival data, especially if Mia can get it set up the way she was thinking to run it using the data structuring work I did. However, right now it looks like nothing is significant for the survival patterns, and similarly not much for emergence. 

Survival plots of Dianthus make it look like light may have had a positive effect on survival in the beginning. I think looking at a continuous spectrum of light might be one way to look at this? 
