---
title: "edivalo-seedlings-toothpicks"
output:
  word_document: default
  html_document: default
---

## Setup and data structuring
Loads libraries, sets up custom theme for ggplot, and reads in data (code not printed)
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
#library(plyr)
library(tidyverse)
library(ggplot2)
library(ggthemes)

# theme for plotting
theme_cw <- function () { 
  theme_bw(base_size=10) %+replace% 
    theme(
      panel.background = element_blank(), 
      plot.background = element_blank(), 
      axis.ticks = element_line(colour = "grey70", size = rel(0.5)),
      panel.grid.minor = element_blank(), 
      panel.grid.major.x = element_blank(),
      legend.background = element_blank(), 
      legend.key = element_blank(),
      strip.background = element_blank(), 
      strip.text=element_text(size=10),
      axis.text=element_text(size=10),
      complete = TRUE
    )
}

# Chhaya's path
mark.seedlings <- read.csv("data/toothpick_seedling_data_2020_all.csv",
                           stringsAsFactors = FALSE)
plot.list <- read.csv("data/plot_list.csv", stringsAsFactors = FALSE)

# Mia's path
#mark.seedlings<-read.csv("C:/Users/mj65tivo/Dropbox/eDiValo-seedlings/Data Entry/toothpick_seedling_data_2020_all.csv", stringsAsFactors = FALSE)
#View(mark.seedlings)
# For Mia: change first column name to month
#colnames(mark.seedlings)[1]<-"month"

# adding separate columns for exclosure, light, and fertilization treatments
mark.seedlings$grazing <- 'sheep'
mark.seedlings$grazing[grep('E', mark.seedlings$plot)] <- 'exclosure'
mark.seedlings$nutrient <- 'unfertilized'
mark.seedlings$nutrient[grep('F', mark.seedlings$plot)] <- 'fertilized'
mark.seedlings$light <- 'control'
mark.seedlings$light[grep('L', mark.seedlings$plot)] <- 'lamps'

# summary of total number by species, plot, and month
mark.summary <- mark.seedlings %>%
    group_by(block, plot, grazing, nutrient, light, species, month) %>%
  dplyr::summarize(
    count = sum(number),
    count.2 = sum(number>0))

# combine plot numbers and treatments together in plot list file
plot.list$plot <- paste(plot.list$plotid, plot.list$treat, sep="")
plot.list$block <- rep(1:10, each = 8)

```

## Visualization plots

### All species through time
Taking a first look at our data, just the abundance counts of the six species through time. Months are listed on the right side (2-February to 7-July). Seedling abundance was generally highest in March-April-May, and the most common species were Plantago lanceolata and Dianthus. 
```{r full, echo=FALSE, warning=FALSE, message=FALSE}
mark.plot <- mark.summary %>% 
  group_by(grazing, nutrient, light, species, month) %>%
  dplyr::summarize(
    plots.present = length(count),
    total.mean = sum(count)/10,
    total.se = sd(count)/sqrt(10))

plantago <- filter(mark.plot, species=="plantago lan")
daucus <- filter(mark.plot, species=="daucus car")
medicago <- filter(mark.plot, species=="medicago fal.")
dianthus <- filter(mark.plot, species=="dianthus car")
galium <- filter(mark.plot, species=="galium alb")
crebis <- filter(mark.plot, species=="crebis bie")

## everything
ggplot(mark.plot, aes(x=nutrient, y=total.mean, 
                     ymin=total.mean-total.se, ymax=total.mean+total.se,
                     color=light, shape=grazing)) +
  facet_grid(month~species) +
  geom_point(size=2, position=position_dodge(0.2)) +
  geom_errorbar(width=0.2, position=position_dodge(0.2)) +
  ylab('Seedling count') +
  xlab('') +
  scale_x_discrete(labels=c("fertilized" = "fert", "unfertilized" = "unfert")) +
  theme_cw() +
  theme(text = element_text(size=10)) +
  scale_color_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'top', legend.text=element_text(size=10))

#ggsave(filename = 'species_time.pdf', width=10, height=8, units='in')

```

Now zooming in on just the most common species to see how they may be responding to treatments.

### April abundance of Plantago
```{r april-plantago, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(filter(plantago, month==4),
       aes(x=nutrient, y=total.mean, 
                     ymin=total.mean-total.se, ymax=total.mean+total.se,
                     color=light)) +
  facet_grid(.~grazing) +
  geom_point(size=4, position=position_dodge(0.5)) +
  geom_errorbar(width=0.2, position=position_dodge(0.5)) +
  ylab('Plantago seedling count') +
  xlab('') +
  theme_cw() +
  theme(text = element_text(size=12)) +
  scale_color_manual(values=c('black','orange'), name='') +
  theme(legend.position = c(0.9, 0.9), legend.text=element_text(size=10))
```


### April abundance of Dianthus

```{r april-dianthus, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(filter(dianthus, month==4),
       aes(x=nutrient, y=total.mean, 
                     ymin=total.mean-total.se, ymax=total.mean+total.se,
                     color=light)) +
  facet_grid(.~grazing) +
  geom_point(size=4, position=position_dodge(0.5)) +
  geom_errorbar(width=0.2, position=position_dodge(0.5)) +
  ylab('Dianthus seedling count') +
  xlab('') +
  theme_cw() +
  theme(text = element_text(size=12)) +
  scale_color_manual(values=c('black','orange'), name='') +
  theme(legend.position = c(0.1, 0.9), legend.text=element_text(size=10))
```

## Emergence
Data re-structuring to look at emergence rather than abundance. We also want to add in zeros for the ones that are missing for the analyses, we're using a complete plot list for this. It is multiplied to have all the species in each plot. 

```{r emergence-data, echo=FALSE, warning=FALSE, message=FALSE}

# for each toothpick, if one or two seedlings emerged
mark.emergence <- mark.seedlings %>%
  group_by(block, plot, grazing, nutrient, light, species, id..1.6.) %>%
  dplyr::summarize(
    emergence.2 = max(number),
    emergence.1 = ifelse(emergence.2 > 0, 1, 0)
    )

# summary of number of seedlings emerged per species per plot
emergence.summary <- mark.emergence %>%
  group_by(block, plot, grazing, nutrient, light, species) %>%
  dplyr::summarise(
    emerged = sum(emergence.2),
    emerged.1 = sum(emergence.1)
  )

# setting up the full data frame with zeros
species.list <- unique(mark.plot$species)
plot.rep <- rep.int(plot.list$plot, length(species.list))
block.rep <- rep.int(plot.list$block, length(species.list))
species.rep <- rep(species.list, each=length(plot.list$plot))

emergence.summary.min <- emergence.summary %>%
  ungroup() %>%
  select(., plot, species, emerged, emerged.1)

emergence.full <- data.frame(plot = plot.rep, species = species.rep, block = block.rep) %>%
  left_join(emergence.summary.min, by = c('plot', 'species'))

# consider turning this into a function?
emergence.full$grazing <- 'sheep'
emergence.full$grazing[grep('E', emergence.full$plot)] <- 'exclosure'
emergence.full$nutrient <- 'unfertilized'
emergence.full$nutrient[grep('F', emergence.full$plot)] <- 'fertilized'
emergence.full$light <- 'control'
emergence.full$light[grep('L', emergence.full$plot)] <- 'lamps'

# changing added 'na' rows to zeros
emergence.full$emerged[is.na(emergence.full$emerged)] <- 0
emergence.full$emerged.1[is.na(emergence.full$emerged.1)] <- 0

```

### Emergence Plot
Plotting emergence by species and treatment
```{r emergence-plot, echo=FALSE, warning=FALSE, message=FALSE}
emergence.plot <- emergence.full %>%
  group_by(grazing, nutrient, light, species) %>%
  dplyr::summarize(
    n = length(emerged),
    emergence.mean = mean(emerged, na.rm = TRUE),
    emergence.se = sd(emerged)/sqrt(n)
  )

ggplot(emergence.plot,
       aes(x=grazing, y=emergence.mean, 
           ymin=emergence.mean-emergence.se, ymax=emergence.mean+emergence.se,
           color=light, shape = nutrient)) +
  facet_wrap(vars(species)) +
  geom_point(size=3, position=position_dodge(0.3)) +
  geom_errorbar(width=0.1, position=position_dodge(0.3)) +
  ylab('Seedling emergence') +
  xlab('') +
  theme_cw() +
  theme(text = element_text(size=10)) +
  scale_color_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'right', legend.text=element_text(size=10))

```
### Emergence Analysis
Starting off with a separate model for each species. We may only be able to get useful models for Plantago and Dianthus, but it's possible that models for Galium and/or Medicago will also converge. We can also explore combining these into models that include species id as an effect (probably having it interact with ).

Probably going to want a negative binomial distribution and/or a zero-inflated model for these emergence counts, similar to the models of the naturally regenerating seedlings. Can approach the model structure and model selection in the same way to keep the methods consistent. 

```{r emergence-analysis}

emerg.pla <- emergence.full %>% filter(species == 'plantago lan')
emerg.dia <- emergence.full %>% filter(species == 'dianthus car')




```


## Next steps
Analyzing emergence for individual species
Analyzing emergence for species combined into a single model?
Structuring and analyzing survival
Incorporating trait data


