---
title: "edivalo-seedlings-toothpicks"
output:
  word_document: default
  html_document: default
---

## Setup and data structuring
Loads libraries, sets up custom theme for ggplot, and reads in data (code not printed)
```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
#library(plyr)
library(tidyverse)
library(ggplot2)
library(ggthemes)

# theme for plotting
theme_cw <- function () { 
  theme_bw(base_size=10) %+replace% 
    theme(
      panel.background = element_blank(), 
      plot.background = element_blank(), 
      axis.ticks = element_line(colour = "grey70", size = rel(0.5)),
      panel.grid.minor = element_blank(), 
      panel.grid.major.x = element_blank(),
      legend.background = element_blank(), 
      legend.key = element_blank(),
      strip.background = element_blank(), 
      strip.text=element_text(size=10),
      axis.text=element_text(size=10),
      complete = TRUE
    )
}

# Chhaya's path
mark.seedlings <- read.csv("data/toothpick_seedling_data_2020_all.csv",
                           stringsAsFactors = FALSE)

# Mia's path
#mark.seedlings<-read.csv("C:/Users/mj65tivo/Dropbox/eDiValo-seedlings/Data Entry/toothpick_seedling_data_2020_all.csv", stringsAsFactors = FALSE)
#View(mark.seedlings)
# For Mia: change first column name to month
#colnames(mark.seedlings)[1]<-"month"

# adding separate columns for exclosure, light, and fertilization treatments
mark.seedlings$grazing <- 'sheep'
mark.seedlings$grazing[grep('E', mark.seedlings$plot)] <- 'exclosure'
mark.seedlings$nutrient <- 'unfertilized'
mark.seedlings$nutrient[grep('F', mark.seedlings$plot)] <- 'fertilized'
mark.seedlings$light <- 'control'
mark.seedlings$light[grep('L', mark.seedlings$plot)] <- 'lamps'

# summary of total number by species, plot, and month
mark.summary <- mark.seedlings %>%
    group_by(block, plot, grazing, nutrient, light, species, month) %>%
  dplyr::summarize(
    count = sum(number),
    count.2 = sum(number>0))

```

## Visualization plots

### All species through time
Taking a first look at our data, just the abundance counts of the six species through time. Months are listed on the right side (2-February to 7-July). Seedling abundance was generally highest in March-April-May, and the most common species were Plantago lanceolata and Dianthus. 
```{r full, echo=FALSE, warning=FALSE, message=FALSE}
mark.plot <- mark.summary %>% 
  group_by(grazing, nutrient, light, species, month) %>%
  dplyr::summarize(
    plots.present = length(count),
    total.mean = sum(count)/10,
    total.se = sd(count)/sqrt(10))

plantago <- filter(mark.plot, species=="plantago lan")
daucus <- filter(mark.plot, species=="daucus car")
medicago <- filter(mark.plot, species=="medicago fal.")
dianthus <- filter(mark.plot, species=="dianthus car")
galium <- filter(mark.plot, species=="galium alb")
crebis <- filter(mark.plot, species=="crebis bie")

## everything
ggplot(mark.plot, aes(x=nutrient, y=total.mean, 
                     ymin=total.mean-total.se, ymax=total.mean+total.se,
                     color=light, shape=grazing)) +
  facet_grid(month~species) +
  geom_point(size=2, position=position_dodge(0.2)) +
  geom_errorbar(width=0.2, position=position_dodge(0.2)) +
  ylab('Seedling count') +
  xlab('') +
  scale_x_discrete(labels=c("fertilized" = "fert", "unfertilized" = "unfert")) +
  theme_cw() +
  theme(text = element_text(size=10)) +
  scale_color_manual(values=c('black','orange'), name='') +
  theme(legend.position = 'top', legend.text=element_text(size=10))

#ggsave(filename = 'species_time.pdf', width=10, height=8, units='in')

```

Now zooming in on just the most common species to see how they may be responding to treatments.

### April abundance of Plantago
```{r april-plantago, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(filter(plantago, month==4),
       aes(x=nutrient, y=total.mean, 
                     ymin=total.mean-total.se, ymax=total.mean+total.se,
                     color=light)) +
  facet_grid(.~grazing) +
  geom_point(size=4, position=position_dodge(0.5)) +
  geom_errorbar(width=0.2, position=position_dodge(0.5)) +
  ylab('Plantago seedling count') +
  xlab('') +
  theme_cw() +
  theme(text = element_text(size=12)) +
  scale_color_manual(values=c('black','orange'), name='') +
  theme(legend.position = c(0.9, 0.9), legend.text=element_text(size=10))
```


### April abundance of Dianthus

```{r april-dianthus, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(filter(dianthus, month==4),
       aes(x=nutrient, y=total.mean, 
                     ymin=total.mean-total.se, ymax=total.mean+total.se,
                     color=light)) +
  facet_grid(.~grazing) +
  geom_point(size=4, position=position_dodge(0.5)) +
  geom_errorbar(width=0.2, position=position_dodge(0.5)) +
  ylab('Dianthus seedling count') +
  xlab('') +
  theme_cw() +
  theme(text = element_text(size=12)) +
  scale_color_manual(values=c('black','orange'), name='') +
  theme(legend.position = c(0.1, 0.9), legend.text=element_text(size=10))
```


## Next steps
Separately analyzing emergence and survival (rather than just abundance) for all species. 
Incorporating trait data. 
